---
title: "Phylogeny S aureus and C jejuni"
author: "Joao Carlos Gomes-Neto"
date: "8/15/2020"
output: html_document
---

Loading the necessary packages 
```{r, include = FALSE}
# Load packages

library(tidyverse)
library(skimr)
library(RColorBrewer)
library(gridExtra)
library(lattice)
library(ggpubr)
library(vegan)
library(reshape2)
library(ggrepel)
library(ggnewscale)
library(forcats)
library(naniar)
library(Rtsne)

# Install ggtree 
#if (!requireNamespace("BiocManager", quietly = TRUE))
  # install.packages("BiocManager")
#BiocManager::install(version = "3.11")

#BiocManager::install("ggtree")

# Load ggtree

library(ggtree)
```

Plot A. C. jejuni plot (phylogeny, ST, and BAPS1)

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("jejuni_phylogeny_gr1.tree")

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('output.roary_output_95_95_1.fastbaps_level_6.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('mlst_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 353, "ST353", 
                ifelse(d1$ST == 45, "ST45",
                ifelse(d1$ST == 982, "ST982",
                ifelse(d1$ST == 48, "ST48",
                ifelse(d1$ST == 50, "ST50",
                ifelse(d1$ST == 8, "ST8",
                ifelse(d1$ST == 806, "ST806",
                ifelse(d1$ST == 459, "ST459",
                       "Other STs"))))))))
                       
# baps

d1$baps1 <- ifelse(d1$baps_1 == 13, "13", 
                   ifelse(d1$baps_1 == 14, "14", 
                          ifelse(d1$baps_1 == 16, "16", 
                                 ifelse(d1$baps_1 == 19, "19", 
                                        ifelse(d1$baps_1 == 22, "22", 
                                               ifelse(d1$baps_1 == 24, "24", 
                       "Other sub-groups"))))))

# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))
d1 <- d1 %>% mutate(baps1 = replace_na(baps1, "Other sub-groups"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps1", "st") 

# Make id the rowname or index

d2 <- column_to_rownames(d2, var = "id") 

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig7a <- gheatmap(t1, d2, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("darkblue", "darkgreen",
                               "cyan", "bisque2", "gray", "coral1",
                                 "orange", "bisque2", "orange",
                               "darksalmon", "chartreuse", "cyan", "darkred", 
                               "darkgreen", "brown1", "darkblue"),
                    breaks = c("13", "14", "16", "19", "22", "24", "Other sub-groups",
                               "ST8", "ST45", "ST48", "ST50", "ST353", "ST459", 
                               "ST806", "ST982", "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle(expression(bold(paste("A. Hierarchical population structure - ", bolditalic(C),". ", bolditalic(jejuni))))) +
  theme(plot.title = element_text(size = 40, face = "bold"), 
        legend.title=element_text(size=24, face = "bold"), 
    legend.text=element_text(size=22))
fig7a
```

Plot B. Select genomes to create phylogeny with all data using Phandango

```{r, include = FALSE}
# enter phylogeny data

tree_group_1 <- read.tree("jejuni_phylogeny_gr1.tree")

# tip label as data frame 

treegenomes <- tree_group_1$tip.label

######################################################################
# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('output.roary_output_95_95_1.fastbaps_level_6.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)


###########################################

# Enter MLST results

mlst <- read_csv('mlst_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 353, "ST353", 
                ifelse(d1$ST == 45, "ST45",
                ifelse(d1$ST == 982, "ST982",
                ifelse(d1$ST == 48, "ST48",
                ifelse(d1$ST == 50, "ST50",
                ifelse(d1$ST == 8, "ST8",
                ifelse(d1$ST == 806, "ST806",
                ifelse(d1$ST == 459, "ST459",
                       "Other STs"))))))))
                       
# baps

d1$baps1 <- ifelse(d1$baps_1 == 13, "13", 
                   ifelse(d1$baps_1 == 14, "14", 
                          ifelse(d1$baps_1 == 16, "16", 
                                 ifelse(d1$baps_1 == 19, "19", 
                                        ifelse(d1$baps_1 == 22, "22", 
                                               ifelse(d1$baps_1 == 24, "24", 
                       "Other sub-groups"))))))

# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))
d1 <- d1 %>% mutate(baps1 = replace_na(baps1, "Other sub-groups"))

#####################################################################

table(d1$st)
table(d1$baps1)

########################################################################

# select columns of interest

data2 <- d1 %>% select(id, baps1, st)

# check for missing values

sum(is.na(data2))

# select columns that are in the tree

data3 <- data2 %>% filter(id %in% treegenomes)

# id to index

data3 <- column_to_rownames(data3, var = "id") 
########################################################################

campypop <- data3

# rename columns

campypop <- rename(campypop, BAPS1 = baps1)
campypop <- rename(campypop, ST = st)

# export training data 
write.csv(campypop, "campypop.csv")

```

Plot C. C. jejuni plot (phylogeny, ST, and BAPS1)
180 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("jejuni_180.tree")

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('jejuni_180_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('mlst_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 353, "ST353", 
                ifelse(d1$ST == 45, "ST45",
                ifelse(d1$ST == 982, "ST982",
                ifelse(d1$ST == 48, "ST48",
                ifelse(d1$ST == 50, "ST50",
                ifelse(d1$ST == 8, "ST8",
                ifelse(d1$ST == 806, "ST806",
                ifelse(d1$ST == 459, "ST459",
                       "Other STs"))))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") %>%
  mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d2 <- column_to_rownames(d2, var = "id") 

# export training data 
popstr180camp <- d2
popstr180camp <- rename(popstr180camp, BAPS1 = baps_1)
popstr180camp <- rename(popstr180camp, ST = st)
write.csv(popstr180camp, "popstr180camp.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig7c <- gheatmap(t1, d2, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("darkblue", "darkgreen",
                               "cyan", "bisque2", "gray", "coral1",
                                 "darkolivegreen", "purple", "orange", "black", "deeppink",
                               "darksalmon", "chartreuse", "cyan", "darkred", 
                               "darkgreen", "brown1", "darkblue", "purple"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10",
                               "ST8", "ST45", "ST48", "ST50", "ST353", "ST459", 
                               "ST806", "ST982", "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("C. Sample size (n = 180 genomes --> 20 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig7c
```

Plot D. C. jejuni plot (phylogeny, ST, and BAPS1)
360 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("jejuni_360.tree")

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('jejuni_360_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('mlst_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 353, "ST353", 
                ifelse(d1$ST == 45, "ST45",
                ifelse(d1$ST == 982, "ST982",
                ifelse(d1$ST == 48, "ST48",
                ifelse(d1$ST == 50, "ST50",
                ifelse(d1$ST == 8, "ST8",
                ifelse(d1$ST == 806, "ST806",
                ifelse(d1$ST == 459, "ST459",
                       "Other STs"))))))))

# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_11)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") %>%
  mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d2 <- column_to_rownames(d2, var = "id") 

# export training data 
popstr360camp <- d2
popstr360camp <- rename(popstr360camp, BAPS1 = baps_1)
popstr360camp <- rename(popstr360camp, ST = st)
write.csv(popstr360camp, "popstr360camp.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig7d <- gheatmap(t1, d2, offset=.0, width=250, colnames = FALSE) +
  scale_fill_manual(values = c("darkblue", "darkgreen",
                               "cyan", "bisque2", "gray", "coral1",
                                 "orange", "purple", "orange",
                               "darksalmon", "chartreuse", "cyan", "darkred", 
                               "darkgreen", "brown1", "darkblue", "purple"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7", "8",
                               "ST8", "ST45", "ST48", "ST50", "ST353", "ST459", 
                               "ST806", "ST982", "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("D. Sample size (n = 360 genomes --> 40 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig7d
```

Plot E. C. jejuni plot (phylogeny, ST, and BAPS1)
540 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("jejuni_540.tree")

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('jejuni_540_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('mlst_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 353, "ST353", 
                ifelse(d1$ST == 45, "ST45",
                ifelse(d1$ST == 982, "ST982",
                ifelse(d1$ST == 48, "ST48",
                ifelse(d1$ST == 50, "ST50",
                ifelse(d1$ST == 8, "ST8",
                ifelse(d1$ST == 806, "ST806",
                ifelse(d1$ST == 459, "ST459",
                       "Other STs"))))))))

# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") %>%
  mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d2 <- column_to_rownames(d2, var = "id") 

# export training data 
popstr540camp <- d2
popstr540camp <- rename(popstr540camp, BAPS1  = baps_1)
popstr540camp <- rename(popstr540camp, ST = st)
write.csv(popstr540camp, "popstr540camp.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig7e <- gheatmap(t1, d2, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("darkblue", "darkgreen",
                               "cyan", "bisque2", "gray", "coral1",
                                 "orange", "purple", "darkseagreen1", "brown", "cornflowerblue",
                               "darksalmon", "chartreuse", "cyan", "darkred", 
                               "darkgreen", "brown1", "darkblue", "purple", "gray"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11",
                               "ST8", "ST45", "ST48", "ST50", "ST353", "ST459", 
                               "ST806", "ST982", "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("E. Sample size (n = 540 genomes --> 60 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig7e
```

Plot F. C. jejuni plot (phylogeny, ST, and BAPS1)
720 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("jejuni_720.tree")

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('jejuni_720_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('mlst_output.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id <- sapply(strsplit(as.character(mlst$FILE),'_'), "[", 1)

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 353, "ST353", 
                ifelse(d1$ST == 45, "ST45",
                ifelse(d1$ST == 982, "ST982",
                ifelse(d1$ST == 48, "ST48",
                ifelse(d1$ST == 50, "ST50",
                ifelse(d1$ST == 8, "ST8",
                ifelse(d1$ST == 806, "ST806",
                ifelse(d1$ST == 459, "ST459",
                       "Other STs"))))))))

# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") %>%
  mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d2 <- column_to_rownames(d2, var = "id") 

# export training data 
popstr720camp <- d2
popstr720camp <- rename(popstr720camp, BAPS1 = baps_1)
popstr720camp <- rename(popstr720camp, ST = st)
write.csv(popstr720camp, "popstr720camp.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig7f <- gheatmap(t1, d2, offset=.0, width=10, colnames = FALSE) +
  scale_fill_manual(values = c("darkblue", "darkgreen",
                               "cyan", "bisque2", "gray", "coral1",
                                 "orange", "purple", "darkseagreen1", "brown", "cornflowerblue", "cornsilk4", "deeppink3",
                               "darksalmon", "chartreuse", "cyan", "darkred", 
                               "darkgreen", "brown1", "darkblue", "purple", "gray"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13",
                               "ST8", "ST45", "ST48", "ST50", "ST353", "ST459", 
                               "ST806", "ST982", "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("F. Sample size (n = 720 genomes --> 80 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig7f
```

#####################################################################################
Plot A. S. aureus plot (phylogeny, ST, and BAPS1)

```{r, include = FALSE}

# Enter the phylogeny for Typhimurium group1

tree_group_1 <- read.tree("saureus_phylogeny_gr1.tree")

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('output.roary_output_1.fastbaps_level_6.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('saureus_st.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 8)
mlst <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 8, "ST8", 
                ifelse(d1$ST == 5, "ST5",
                ifelse(d1$ST == 105, "ST105",
                ifelse(d1$ST == 398, "ST398",
                ifelse(d1$ST == 609, "ST609",
                ifelse(d1$ST == 45, "ST45",
                       "Other STs"))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") %>%
                    mutate(baps_1 = as.factor(baps_1))

# Make id the rowname or index

d2 <- column_to_rownames(d2, var = "id") 

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig8a <- gheatmap(t1, d2, offset=.0, width=150, colnames = FALSE) +
  scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "darkseagreen3", "darkcyan", "gray",
                               "brown", "black", "darkblue", "purple", 
                               "orange"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7",
                               "ST5", "ST8", "ST45", "ST105", "ST398", "ST609", 
                               "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle(expression(bold(paste("A. Hierarchical population structure - ", bolditalic(S),". ", bolditalic(aureus))))) +
  theme(plot.title = element_text(size = 40, face = "bold"), 
        legend.title=element_text(size=24, face = "bold"), 
    legend.text=element_text(size=22))
fig8a
```

Plot B. S. aureus plot (phylogeny, ST, and BAPS1)

```{r, include = FALSE}

# Enter the phylogeny for Typhimurium group1

tree_group_1 <- read.tree("saureus_phylogeny_gr1.tree")

# tip label as data frame 

treegenomes <- tree_group_1$tip.label

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('output.roary_output_1.fastbaps_level_6.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('saureus_st.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 8)
mlst <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 8, "ST8", 
                ifelse(d1$ST == 5, "ST5",
                ifelse(d1$ST == 105, "ST105",
                ifelse(d1$ST == 398, "ST398",
                ifelse(d1$ST == 609, "ST609",
                ifelse(d1$ST == 45, "ST45",
                       "Other STs"))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") %>%
                    mutate(baps_1 = as.factor(baps_1))

# select columns that are in the tree

data3 <- d2 %>% filter(id %in% treegenomes)

# id to index

data3 <- column_to_rownames(data3, var = "id") 
########################################################################

saureuspop <- data3

# rename columns

saureuspop <- rename(saureuspop, BAPS1 = baps_1)
saureuspop <- rename(saureuspop, ST = st)

# export training data 
write.csv(saureuspop, "saureuspop.csv")
```

Plot C. S. aureus plot (phylogeny, ST, and BAPS1)
140 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("saureus_140.tree")

# tip label as data frame 

treegenomes <- tree_group_1$tip.label

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('saureus_140_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('saureus_st.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 8)
mlst <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 8, "ST8", 
                ifelse(d1$ST == 5, "ST5",
                ifelse(d1$ST == 105, "ST105",
                ifelse(d1$ST == 398, "ST398",
                ifelse(d1$ST == 609, "ST609",
                ifelse(d1$ST == 45, "ST45",
                       "Other STs"))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") 

# filter rows that are included in the tree

data3 <- d2 %>% filter(id %in% treegenomes)

# Make id the rowname or index

data3 <- column_to_rownames(data3, var = "id") 

# export training data 
popstr140sa <- data3
popstr140sa <- rename(popstr140sa, BAPS1 = baps_1)
popstr140sa <- rename(popstr140sa, ST = st)
write.csv(popstr140sa, "popstr140sa.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-10, NA)


fig8c <- gheatmap(t1, data3, offset=.0, width=250, colnames = FALSE) +
scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "darkseagreen3", "darkcyan", "gray",
                               "brown", "black", "darkblue", "purple"),
                    breaks = c("1", "2", "3", "4", "5", "6",
                               "ST5", "ST8", "ST45", "ST105", "ST398", "ST609", 
                               "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("C. Sample size (n = 137 genomes --> ~ 20 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig8c
```

Plot D. S. aureus plot (phylogeny, ST, and BAPS1)
350 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("saureus_350.tree")

# tip label as data frame 

treegenomes <- tree_group_1$tip.label

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('saureus_350_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('saureus_st.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 8)
mlst <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 8, "ST8", 
                ifelse(d1$ST == 5, "ST5",
                ifelse(d1$ST == 105, "ST105",
                ifelse(d1$ST == 398, "ST398",
                ifelse(d1$ST == 609, "ST609",
                ifelse(d1$ST == 45, "ST45",
                       "Other STs"))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") 

# filter rows that are included in the tree

data3 <- d2 %>% filter(id %in% treegenomes)

# Make id the rowname or index

data3 <- column_to_rownames(data3, var = "id") 

# export training data 
popstr350sa <- data3
popstr350sa <- rename(popstr350sa, BAPS1 = baps_1)
popstr350sa <- rename(popstr350sa, ST = st)
write.csv(popstr350sa, "popstr350sa.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig8d <- gheatmap(t1, data3, offset=.0, width=10, colnames = FALSE) +
scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "darkseagreen3", "darkcyan", "gray",
                               "brown", "black", "darkblue", "purple"),
                    breaks = c("1", "2", "3", "4", "5", "6",
                               "ST5", "ST8", "ST45", "ST105", "ST398", "ST609", 
                               "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("D. Sample size (n = 349 genomes --> ~ 50 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig8d
```

Plot E. S. aureus plot (phylogeny, ST, and BAPS1)
560 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("saureus_560.tree")

# tip label as data frame 

treegenomes <- tree_group_1$tip.label

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('saureus_560_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('saureus_st.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 8)
mlst <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 8, "ST8", 
                ifelse(d1$ST == 5, "ST5",
                ifelse(d1$ST == 105, "ST105",
                ifelse(d1$ST == 398, "ST398",
                ifelse(d1$ST == 609, "ST609",
                ifelse(d1$ST == 45, "ST45",
                       "Other STs"))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") 

# filter rows that are included in the tree

data3 <- d2 %>% filter(id %in% treegenomes)

# Make id the rowname or index

data3 <- column_to_rownames(data3, var = "id") 

# export training data 
popstr560sa <- data3
popstr560sa <- rename(popstr560sa, BAPS1 = baps_1)
popstr560sa <- rename(popstr560sa, ST = st)
write.csv(popstr560sa, "popstr560sa.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig8e <- gheatmap(t1, data3, offset=.0, width=250, colnames = FALSE) +
scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "darkseagreen3", "darkcyan", "gray",
                               "brown", "black", "darkblue", "purple", "orange"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7",
                               "ST5", "ST8", "ST45", "ST105", "ST398", "ST609", 
                               "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("E. Sample size (n = 558 genomes --> ~ 80 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig8e
```

Plot F. S. aureus plot (phylogeny, ST, and BAPS1)
770 genomes

```{r, include = FALSE}

# Enter the phylogeny 

tree_group_1 <- read.tree("saureus_770.tree")

# tip label as data frame 

treegenomes <- tree_group_1$tip.label

#############################################

# Enter Baps 1-6 for group1

baps_group_1 <- read_csv('saureus_770_fastbaps.csv')

# Changing column names 

colnames(baps_group_1)[1:2] <- c("id", "baps_1") 

# Select colunms id and baps_1
b1 <- baps_group_1 %>% 
              select(id, baps_1)

# QC the data

skim(b1)

###########################################

# Enter MLST results

mlst <- read_csv('saureus_st.csv')

# QC the data

skim(mlst)

# Generate the id column 

mlst$id2 <- sapply(strsplit(as.character(mlst$FILE),'/'), "[", 8)
mlst <- mlst %>% separate(id2, c("id", "day"), sep = "\\.")

# Select columns of interest

s1 <- mlst %>%
          select(id, ST) %>%
            mutate_all(na_if, "-") %>%
              mutate_all(na_if, "?")

# QC the data

skim(s1)

#####################################################################

# Merge BAPS, MLST and SISTR results 

## Merge baps and mlst first

d1 <- left_join(b1, s1, on = "id") # merge mlst and baps

# st

d1$st <- ifelse(d1$ST == 8, "ST8", 
                ifelse(d1$ST == 5, "ST5",
                ifelse(d1$ST == 105, "ST105",
                ifelse(d1$ST == 398, "ST398",
                ifelse(d1$ST == 609, "ST609",
                ifelse(d1$ST == 45, "ST45",
                       "Other STs"))))))
                       
# Replace NA in the ST column

d1 <- d1 %>% mutate(st = replace_na(st, "Other STs"))

table(d1$st)
table(d1$baps_1)

########################################################################

# Plot the phylogeny

# Select the columns needed from d2

d2 <- d1 %>% 
                    select("id", "baps_1", "st") 

# filter rows that are included in the tree

data3 <- d2 %>% filter(id %in% treegenomes)

# Make id the rowname or index

data3 <- column_to_rownames(data3, var = "id") 

# export training data 
popstr770sa <- data3
popstr770sa <- rename(popstr770sa, BAPS1 = baps_1)
popstr770sa <- rename(popstr770sa, ST = st)
write.csv(popstr770sa, "popstr770sa.csv")

# Create the tree

t1 <- ggtree(tree_group_1, layout = "circular") + xlim(-50, NA)


fig8f <- gheatmap(t1, data3, offset=.0, width=250, colnames = FALSE) +
scale_fill_manual(values = c("gray", "darkblue",
                               "cornflowerblue", "coral", "darkgreen", "red",
                                 "darkseagreen3", "darkcyan", "gray",
                               "brown", "black", "darkblue", "purple", "orange"),
                    breaks = c("1", "2", "3", "4", "5", "6", "7",
                               "ST5", "ST8", "ST45", "ST105", "ST398", "ST609", 
                               "Other STs"),
                    name= "BAPS1 (innermost) -> ST (outermost)") +
  ggtitle("F. Sample size (n = 766 genomes --> ~ 110 per ST)") +
  theme(plot.title = element_text(size = 50, face = "bold"), 
        legend.title=element_text(size=35, face = "bold"), 
    legend.text=element_text(size=30))
fig8f
```
